---
title: "education skills job"
author: "Shangjun Jiang"
date: "29/09/2017"
output: html_document
---



1. total by jobs, detaield education by year
TOTAL NUMBER

big effect

by aboslute numbers
range calculations.
within job types

how many people in each type of jobs

certain jobs really change

percentage data, range really high
divide by bachelor, above and below,high school

jobs by occupations

numbers for all occupations. 

change within overall occupations 

facet slope graph 

#### separate percentage and numebrs. 
join employment number and percentage. education by detailed occupations 



2. employment projection.
2012 total numbers and projections. 2012-2022;  2014-2024
line chart. 

absolute numbers.
outliers. 

jobs emerging. education-under-class
high education
any go other directions. 


3. by dot and lines
year, wage, degrees color
year, unemployment, degrees color 
calculate, changes, percentage, 




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(MASS)
library(hexbin)
library(corrplot)
library(car)
library(ggplot2)
require(methods)
require("leaps")
library(tidyr)
library(reshape2)
library(ridge)
require(lars)
library(descr)
library(stringr)
library(readr)
library(readxl)
library(tidyverse)
#library(plyr)
library(dplyr)
#setwd("~/Desktop/week7")
#save.image("~/Desktop/week7/week7.RData")
#setwd("~/Desktop/labor data")

setwd("~/Desktop/jenny-jiang")
```





```{r}
########################  EDUCATION by detailed occupations ########################  
emp_edu_1015<- read.csv("emp_edu_1015.csv")
emp_edu_1015$X <- NULL
emp_edu_1015$year <- as.factor(emp_edu_1015$year)
### wide to long format 

melt1015 <- melt(emp_edu_1015, 
     measure.vars = c("less.hs", "high.school", "associate", "some.college..no.degree", "bachelor", "master", "doctor.professional"), 
     variable.name = "degrees", 
     value.name = "number")

melt1015$edu <- melt1015$degrees
melt1015$year <- as.factor(melt1015$year)

# college +
levels(melt1015$edu)[levels(melt1015$edu)=="bachelor"] <- "college+"
levels(melt1015$edu)[levels(melt1015$edu)=="doctor.professional"] <- "college+"
levels(melt1015$edu)[levels(melt1015$edu)=="master"] <- "college+"
# hs - college
levels(melt1015$edu)[levels(melt1015$edu)=="some.college..no.degree"] <- "hs-college"
levels(melt1015$edu)[levels(melt1015$edu)=="associate"] <- "hs-college"

# hs -
levels(melt1015$edu)[levels(melt1015$edu)=="high.school"] <- "hs-"
levels(melt1015$edu)[levels(melt1015$edu)=="less.hs"] <- "hs-"
levels(melt1015$edu)


#write.csv(melt1015, "long_edu_employ_1015.csv")

#4830 obs 9 variables
#str(melt1015)

total_summary <- melt1015 %>%
  group_by(year, edu) %>%
  summarize(sum=sum(number, na.rm=T))


ggplot(total_summary, 
           aes(x=year, y=sum, group=edu, color=edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total by edu",
           x = "Year",
           y = "total sum")



#2070 obs, 4 variables
edu_occ_summary <- melt1015 %>%
  group_by(year, edu, employment, code) %>%
  summarize(sum=sum(number, na.rm=T))


ggplot(edu_occ_summary, 
           aes(x=year, y=sum, group=edu, color=edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total by edu",
           x = "Year",
           y = "total sum")


####### computer and math occupations
CS_related <- edu_occ_summary[grepl("15-", edu_occ_summary$code),]


CS_related_summary <- CS_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(CS_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in CS-related",
           x = "Year",
           y = "total number")

####### business and financial occupations
biz_related <- edu_occ_summary[grepl("13-", edu_occ_summary$code),]


biz_related_summary <- biz_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(biz_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in Business-related",
           x = "Year",
           y = "total number")


####### arts, media occupations
art_related <- edu_occ_summary[grepl("27-", edu_occ_summary$code),]


art_related_summary <- art_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(art_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in Arts-related",
           x = "Year",
           y = "total number")

####### healthcare occupations
health1_related <- edu_occ_summary[grepl("29-", edu_occ_summary$code),] 
health2_related <- edu_occ_summary[grepl("31-", edu_occ_summary$code),] 

health_related <- full_join(health1_related, health2_related) 


health_related_summary <- health_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(health_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in health-related",
           x = "Year",
           y = "total number")

####### sales occupations
sales_related <- edu_occ_summary[grepl("41-", edu_occ_summary$code),] 

sales_related_summary <- sales_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(sales_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in sales-related",
           x = "Year",
           y = "total number")

####### office occupations
office_related <- edu_occ_summary[grepl("43-", edu_occ_summary$code),] 


office_related_summary <- office_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(office_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in office-related",
           x = "Year",
           y = "total number")




####### service occupations
service1_related <- edu_occ_summary[grepl("35-", edu_occ_summary$code),] 
service2_related <- edu_occ_summary[grepl("37-", edu_occ_summary$code),] 

service_related <- full_join(service1_related, service2_related) 


service_related_summary <- service_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(service_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in service-related",
           x = "Year",
           y = "total number")


####### transportation, production occupations
production1_related <- edu_occ_summary[grepl("53-", edu_occ_summary$code),] 
production2_related <- edu_occ_summary[grepl("51-", edu_occ_summary$code),] 
production3_related <- edu_occ_summary[grepl("47-", edu_occ_summary$code),] 
production4_related <- edu_occ_summary[grepl("49-", edu_occ_summary$code),] 

production12_related <- full_join(production1_related, production2_related) 
production34_related <- full_join(production3_related, production4_related) 
production_related <- full_join(production34_related, production12_related) 

production_related_summary <- production_related %>%
  group_by(year, edu) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(production_related_summary, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total number in production-related",
           x = "Year",
           y = "total number")


```



```{r}
setwd("~/Desktop/jenny-jiang")
########################  EDUCATION by detailed occupations ########################  
edu1015long <- read.csv("edu1015long.csv")
edu1015long$year <- as.factor(edu1015long$year)
#str(edu1015long)
edu1015long$edu <- edu1015long$degrees
#levels(edu1015long$degrees)

# college +
levels(edu1015long$edu)[levels(edu1015long$edu)=="bachelor"] <- "college+"
levels(edu1015long$edu)[levels(edu1015long$edu)=="doctor/professional"] <- "college+"
levels(edu1015long$edu)[levels(edu1015long$edu)=="master"] <- "college+"
# hs - college
levels(edu1015long$edu)[levels(edu1015long$edu)=="some college, no degree"] <- "hs+"
levels(edu1015long$edu)[levels(edu1015long$edu)=="associate"] <- "hs+"

# hs -
levels(edu1015long$edu)[levels(edu1015long$edu)=="high school"] <- "hs-"
levels(edu1015long$edu)[levels(edu1015long$edu)=="less hs"] <- "hs-"
levels(edu1015long$edu)

#str(edu1015long)


#7389
edu_occ_summary <- edu1015long %>%
  group_by(year, edu, code, year) %>%
  summarize(sum=sum(number, na.rm=T))



check1 <- edu_occ_summary[edu_occ_summary$code == "00-0000",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total",
           x = "Year",
           y = "Percentage")



########################  EDUCATION AND UNEMPLOYMENT RATE ########################  
unemploy1216 <- read.csv("unemploy1216.csv")
unemploy1216$education <- as.factor(unemploy1216$education)
unemploy1216$year <- as.factor(unemploy1216$year)
unemploy1216$X <- NULL
#str(unemploy1216)


######################## EDUCATION AND TRAINING ASSIGNMENTS BY DETAILED OCCUPATION ##################
edutrain1014 <- read.csv("edutrain1014.csv")
edutrain1014$year <- as.factor(edutrain1014$year)
edutrain1014$X <- NULL
#str(edutrain1014)

########################  EMPLOYMENT PROJECTION ########################
employ_proj1214 <- read.csv("employ_proj1214.csv")
employ_proj1214$year <- as.factor(employ_proj1214$year)
employ_proj1214$annual_wage <- as.numeric(employ_proj1214$annual_wage)
employ_proj1214$X <- NULL
#str(employ_proj1214)


############### OCCUPATIONS THAT NEED MORE EDUCATION FOR ENTRY ARE PROJECTED TO GROW FASTER ########################
edu_level1014 <- read.csv("edu_level1014.csv")
edu_level1014$X <- NULL
edu_level1014$year <- as.factor(edu_level1014$year)

job_training1012 <- read.csv("job_training1012.csv")
job_training1012$X <- NULL
job_training1012$year <- as.factor(job_training1012$year)

work_experience1012 <- read.csv("work_experience1012.csv")
work_experience1012$X <- NULL
work_experience1012$year <- as.factor(work_experience1012$year)

#str(edu_level1014)
#str(job_training1012)
#str(work_experience1012)


```


# by college level
```{r}

edu_summary <- edu_occ_summary %>%
  group_by(year, edu, year) %>%
  summarize(sum=sum(sum, na.rm=T))

ggplot(edu_summary , 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total by edu",
           x = "Year",
           y = "total sum")


```



# Manager level and non-manager 

```{r}

edu_managers <- edu1015long[grep('managers', edu1015long$employment), ]

#546
edu_managers <- edu1015long %>%
  filter(str_detect(employment, "manager"))

#16,695
edu_nonmanagers <- edu1015long %>%
  filter(!str_detect(employment, "manager"))

#234
edu_manager_summary <- edu_managers %>%
  group_by(year, edu, code) %>%
  summarize(sum=sum(number, na.rm=T))

check1 <- edu_manager_summary[edu_manager_summary$code == "2648679",]

edu_manager_summary1 <- edu_managers %>%
  group_by(year, edu, number) %>%
  summarize(sum=sum(number, na.rm=T))


ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total",
           x = "Year",
           y = "Percentage")



```


##### different occupations

```{r}
check1 <- edu_occ_summary[edu_occ_summary$code == "00-0000",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total",
           x = "Year",
           y = "Percentage")


check1 <- edu_occ_summary[edu_occ_summary$code == "11-1021",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_line() +
      geom_point() +
      labs(title = "General and operations managers",
           x = "Year",
           y = "Percentage")


check1 <- edu_occ_summary[edu_occ_summary$code == "2605581",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_line() +
      geom_point() +
      labs(title = "Education administrators, postsecondary",
           x = "Year",
           y = "Percentage")


check1 <- edu_occ_summary[edu_occ_summary$code == "2619460",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_line() +
      geom_point() +
      labs(title = "Gaming managers",
           x = "Year",
           y = "Percentage")


check1 <- edu_occ_summary[edu_occ_summary$code == "11-1021",]

ggplot(check1, 
           aes(x=year, y=sum, group=edu, color =edu)) +
      geom_line() +
      geom_point() +
      labs(title = "occupation",
           x = "Year",
           y = "Percentage")



```










## EDUCATION AND UNEMPLOYMENT RATE
```{r}
ggplot(unemploy1216, aes(x=education, y=weekly_wage, fill=year))+
  geom_bar(width = 1, stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


## EMPLOYMENT PROJECTION


```{r}
#employ_proj1214


```



##EDUCATION AND TRAINING ASSIGNMENTS BY DETAILED OCCUPATION
this is more of reference/ code book. 
 
```{r}

#edutrain1014


```







```{r}
ggplot(edu_occ_summary, 
           aes(x=year, y=sum, group=edu, color=edu)) +
      geom_point() +
      geom_line() +
      labs(title = "total by edu",
           x = "Year",
           y = "total sum")

```



