---
title: "other codes"
author: "Shangjun Jiang"
date: "12/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}


```{r}
food_stamps_column <- ggplot(total_summary, aes(x = edu, y = sum)) + 
  xlab("Year") +
  ylab("total number of employees") +
  theme_minimal(base_size = 14, base_family = "Georgia") + 
  geom_bar_interactive(stat = "identity", 
         color = "#888888", 
         fill = "#CCCCCC", 
         alpha = 0.5,
         aes(tooltip = paste0("number: ",sum),
             data_id = edu)) 

# make interactive version of the chart
food_stamps_column_interactive <- ggiraph(code = print(food_stamps_column), 
                                   height_svg = 4,
                                   hover_css = "fill-opacity:1;stroke:red",
                                   tooltip_extra_css = "background-color:#f0f0f0;color:black;padding:5px")

food_stamps_column_interactive

mtcars$names <- rownames(mtcars)
scatterD3(data = mtcars, x = wt, y = mpg, lab = names,
          col_var = cyl, symbol_var = am,
          xlab = "Weight", ylab = "Mpg", col_lab = "Cylinders",
          symbol_lab = "Manual transmission")

melt1015$names <- rownames(melt1015)


          
scatterD3(data = melt1015, x = year, y = number, 
          col_var = degrees, symbol_var = edu,
          size_var = median_salary, 
          size_range = c(10,1000), point_opacity = 0.7)



computer_info %>%
  mjs_plot(x=degrees, y=Total, width=600, height=500) %>%
  mjs_point(color_accessor=edu,
            x_rug=TRUE, y_rug=TRUE,
            color_type="category",
            color_range=brewer.pal(n=11, name="RdBu")[c(1, 5, 11)],
            least_squares=TRUE)



billboarder() %>% 
  bb_scatterplot(data = melt1015) %>% 
  bb_linechart(data = data.frame(lowess(melt1015)), x = "median_salary") %>% 
  bb_data(types = list(dist = "scatter", y = "Total"))

    cols <- brewer.pal(7, "Set1")
    
    
    hchart(cleaned1, "line", hcaes(x = "year", y = "sum", group = "edu")) %>%
      hc_colors(cols) %>%
      hc_chart(style = list(fontFamily = "Georgia",
                            fontWeight = "bold")) %>%
      hc_title(text = input$question,
               margin = 20, align = "center",
               style = list(useHTML = TRUE)) %>% 
      hc_xAxis(title = list(text="Year")) %>%
      hc_yAxis(title = list(text="Total Number of Workers")) %>%
      hc_plotOptions(series = list(marker = list(enabled = TRUE, symbol = "circle")))
    
    
    
    
```




```

